---
title: "Financia Engineering - Crude oil prices"
author: "AyşegülBinbaş" 
date: "04 02 2022"
output: word_document
---

```{r}

library(dplyr)
library(magrittr)
library(readxl)
library(TSA)
library(forecast)



oil <- read_excel("Crudeoilprices.xlsx")
oil$Date<-as.Date(oil$Date,format="%d.%m.%y")
oil<-oil%>%mutate(Crude.Oil.Price=as.numeric(gsub(",",".",Crude.Oil.Price)))




S0 <- 0.15   
mu <- mean(exp(oil$Crude.Oil.Price))
sd <- sd(exp(oil$Crude.Oil.Price))
T <- 0.5



(mean <- log(S0) + (mu-sd^2/2)*T)
(var   <- sd^2*T)
(stdev <- sqrt(var))

(CI <- c(mean-1.96*stdev, mean+1.96*stdev))

vol_percent = sd(exp(oil$Crude.Oil.Price)) / mean(exp(oil$Crude.Oil.Price))
vol_percent 


monthly_sigma <- sd * sqrt(252/12)  
monthly_sigma

BlackScholes <- function(S, K, r, T, sig, type){
  
  if(type=="C"){
    d1 <- (log(S/K) + (r + sig^2/2)*T) / (sig*sqrt(T))
    d2 <- d1 - sig*sqrt(T)
    
    value <- S*pnorm(d1) - K*exp(-r*T)*pnorm(d2)
    return(value)}
  
  if(type=="P"){
    d1 <- (log(S/K) + (r + sig^2/2)*T) / (sig*sqrt(T))
    d2 <- d1 - sig*sqrt(T)
    
    value <-  (K*exp(-r*T)*pnorm(-d2) - S*pnorm(-d1))
    return(value)}
}


#This function first calculates the d1 and d2 parameters required for the Black-Scholes model and then uses pnorm() command of R which simulates a cumulative normal distribution.
# In order to use the BlackScholes function to value a call and a put option, we can run the following lines:


put <- BlackScholes(0.15,0.3,0.08,0.5,0.48,"P")
put
Put <- function(S, K, r, T, sigma) {
  d1  <-  (log(S/K) + (r + sigma^2/2)*T) / (sigma*sqrt(T))
  d2  <-  d1 - sigma*sqrt(T)
  -S * pnorm(-d1) + K*exp(-r*T)*pnorm(-d2)
}

```




Assumptions of the Black-Scholes Model
Lognormal distribution: It was concluded that this assumption was violated because the converted data did not follow a normal distribution with outliers.

No dividends: We assume that the Black-Scholes model does not require any initial payment or returns.

Expiration date: As we tried to put option price with the maturity duration of 6 months after the end date of our data set, this assumption already exists.

Random walk: We checked this assumption and we showed below.

Frictionless market: As our model does not include the costs of commission and brokerage, the Frictionless market assumption is applied for our analysis.

Risk-free interest rate: The nominal interest rate is 8% is given to us in the question.

Normal distribution: Stock returns are not normally distributed. We assume that the volatility of the market is constant over time.

No arbitrage: There is no arbitrage because of the free arbitrage in part a.

```{r}

library(dplyr)
library(magrittr)
library(readxl)
library(TSA)
library(forecast)







#Checking The Stationarity Assumption
CrudeOilPrice<-oil$Crude.Oil.Price
plot(CrudeOilPrice, type = "l", lwd = 2.5, xlab = "Time", ylab = "Crude Spot Price", col = 1, main = "Historical Spot Price for Brent Fuel")


#When we look at the graph, 
#ıt can be seen that there exist some up and downs ,and there is no constant mean aroud 0.
# Hence, ıt can be saıd that the data is not stationary.

hist(CrudeOilPrice, main='Crude Oil Spot Prices', xlab='Crude Oil Price', col='darkblue', breaks = 1000, border="red", freq=F)
#lines(density(CrudeOilPrice))

library(pdR)
library(tseries)



ndiffs(CrudeOilPrice)
#Taking 1st time difference is enough.
#nsdiffs(CrudeOilPrice)
#Non seasonal data
# To solve the non stationary problem,we will difference of our data. 
difcrudeoilprice = diff(CrudeOilPrice)
head(difcrudeoilprice)

plot(difcrudeoilprice, type = "l", lwd = 2.5, xlab = "Time", ylab = "CrudeLog Returns", col = 1, main = "Brent Crude Oil Log Returns")

#As can be seen from the graph that ıt seem stationart , but there are some outliers.To be sure we can apply some formal tests.

kpss.test(difcrudeoilprice,null=c("Level")) #to check stationary or not

#p-value greater than printed p-value,so we cannot reject H0.
# That means we have enough evidence to claim that the process is stationary.
#hence, that is the smoothed log-returns or smoothed increments, ln(St)-ln(st-1) of the real oil prices are stationary.

#ADF test: to cleck roots

adf.test(difcrudeoilprice)
# p-value = 0.0001.Therefore, These results suggest the rejection of the
#null hypothesis which is implying the stationarity of the standardized increments of the smoothed real oil prices.

#For ACF:
p1 <- ggAcf(difcrudeoilprice,main="ACF of Crude Oil Prices")

# For PACF :
p2 <-ggPacf(difcrudeoilprice,main="PACF of Crude Oil Prices")

library(gridExtra)

grid.arrange(p1,p2,ncol=2)


#ACF look like stationary,but there are some seasonal patterns.According to PACF we can say that lag 15 is significant.


#Checking The Normality Assumption

#In order to check the normality of the standardized log-returns we perform the Jarque-Bera test in which
# Ho:data standard normal distribution
#Ha:data is not standard normal distribution

library(tseries)


jarque.bera.test(difcrudeoilprice)

# p-value < 2.2e-16 that is less than alpha so, data satisfied the normality assumption.

hist(difcrudeoilprice, freq=FALSE) 

# It is follow normal distribution,according to test.However,there exist some outliers.
#Wecan also,clean this outliers
cleaned_difcrudeoilprice=tsclean(difcrudeoilprice)
hist(cleaned_difcrudeoilprice, freq=FALSE) 
#Afher using ts clean fns , the shape of the histogram more clear.

#Checking The Dependence Assumption
#checkresiduals(cleaned_difcrudeoilprice)

#We can also use correlogram of the autocorrelation and partial autocorrelation
#there is dependence in the data set.


# After checking all assumptions, we replaced the given values and found the put option to be 0.1389608.

```

